# Four Step Process Notes
```{r setup, echo = FALSE}
# load weightloss4 file
weightloss <- read.csv("weightloss.csv", header = TRUE)
attach(weightloss)
```

### Produce cooperative boxplots
```{r}
boxplot(WeightLoss ~ Group)
```

### Produce cooperative dotplots
```{r}
stripchart(WeightLoss ~ Group, method = "stack", pch = 20, cex = 1.5)
```

Find the n, mean, and sd for each condition

- We could find each item individually
- Note the double equals sign...
  - A single equals sign (=) is an arithmetic operator
  - A double equals sign (==) is a logical operator (tests True/False)
```{r}
mu_control <- mean(WeightLoss[Group == "Control"])
```

#### OR use tapply() function to see both conditions at once
```{r}
tapply(WeightLoss, Group, length)
tapply(WeightLoss, Group, mean)
tapply(WeightLoss, Group, sd)
```

### Create a table for the n, mean, and sd according to condition
```{r}
n <- tapply(WeightLoss, Group, length)
mean <- tapply(WeightLoss, Group, mean)
SD <- tapply(WeightLoss, Group, sd)
cbind(n, mean, SD)
```

### Round table output to 2 decimal places
```{r}
round(cbind(n, mean, SD), 2)
```

### T-test
```{r}
t.test(WeightLoss ~ Group)
```

### Look at Residuals (many steps to follow)
Treat Group condition as numeric (Control = 0 and Incentive = 1)

Place the numeric condition in a new column called GroupNumeric
```{r}
GroupNumeric <- as.numeric(Group == "Incentive")
mu_control <- mean(WeightLoss[GroupNumeric == 0])
mu_incentive <- mean(WeightLoss[GroupNumeric == 1])
```
Now we're taking steps to make a column that will contain the correct mean for the condition of Group variable

Notice how the math works:

- For all people in the control group, where `GroupNumeric = 0`, the red group below will = 0, and we will be left with the green group (mu_control * 1)
- For all people in the control group, where `GroupNumeric = 1`, the green group below will = 0, and we will be left with the red group (mu_incentive * 1)
```{r}
Fit <- mu_control * (1 - GroupNumeric) + mu_incentive * GroupNumeric
print(Fit)
```

### Find Residuals (Difference from observed values and mean)
```{r}
Resid <- WeightLoss - Fit
# Put them all together
cbind(WeightLoss, Fit, Resid)
```

We already have t-statistic, but we can also calculate it manually

Compare each result with the computer output from the t-test
```{r}
sdc <- sd(WeightLoss[Group == "Control"])
sdi <- sd(WeightLoss[Group == "Incentive"])
nc <- length(WeightLoss[Group == "Control"])
ni <- length(WeightLoss[Group == "Incentive"])
t.stat <- (mu_control - mu_incentive) / sqrt(sdc^2 / nc + sdi^2 / ni)
```

### Check the Residuals

The qqnorm() function computes a normal probability plot

The qqline() function adds a reference line to this plot to help assess the straightness of the normal plot
```{r}
qqnorm(Resid[Group == "Control"], ylab = "Control Residuals")
qqline(Resid[Group == "Control"])
qqnorm(Resid[Group == "Incentive"], ylab = "Incentive Residuals")
qqline(Resid[Group == "Incentive"])
```

### Effect size

We know from the p-value that the difference of 11.76 pounds is too big to be due to the random assignment of subjects to groups (chance).

We know from the margin of error that with 95% confidence, the incentive is worth between 5.5 and 18.1 pounds.

What we don't know yet is how much of a difference 11.76 pounds actually matters.

The effect size is the ratio of observed differences (differences in means) to the standard deviation for person-to-person differences (pooled sd)
`=(abs(mu1 - mu2))/(pooled sd)`
```{r}
# numerator
num <- abs(mu_control - mu_incentive)
# denominator
den <- sqrt(((nc - 1) * sdc^2 + (ni - 1) * sdi^2) / (nc + ni - 2))
effect <- num / den
```

This tells us that the effect of the incentive is 127% as big as the typical size of the difference between you and an average person.