# R Basics Quiz 09/06/24
Directions: In your R Script window, copy and paste the items below.

Under each item, write the appropriate code.

Copy and paste the contents of your script window into a word file.

Submit.


### Load and attach the BirdNest data
```{r}
birdnest <- read.csv("BirdNest.csv", header = TRUE)
attach(birdnest)
```
### Look at the entry in the 28 th row, 4 th column
```{r}
birdnest[28, 4]
```
### Look at all entries for the 28 th row
```{r}
birdnest[28, ]
```
### Look at all entries for the 4 th column
```{r}
birdnest[, 4]
```
### Create a new column called “TotalCare” which adds the entries for Incubate and Nestling
```{r}
birdnest$TotalCare <- birdnest$Incubate + birdnest$Nestling
```
### Sort the dataset according to the TotalCare column in ascending order
```{r}
birdnest <- birdnest[order(birdnest$TotalCare), ]
```
### Find the mean of the TotalCare column
```{r}
mean(birdnest$TotalCare)
```
### Find the median of the TotalCare column
```{r}
median(birdnest$TotalCare)
```
### Find the standard deviation of the TotalCare column
```{r}
sd(birdnest$TotalCare)
```
### Find the variance of the TotalCare column
```{r}
var(birdnest$TotalCare)
```
### Find the 5-number summary of the TotalCare column
```{r}
summary(birdnest$TotalCare)
```
### Create a stemplot for TotalCare
```{r}
stem(birdnest$TotalCare)
```
### Create a boxplot for TotalCare
```{r}
boxplot(birdnest$TotalCare)
```
### Create a histogram for TotalCare
```{r}
hist(birdnest$TotalCare)
```
### Adjust histogram so that there are 5 – 8 bins
```{r}
hist(birdnest$TotalCare, breaks = 6)
```

### Create a scatterplot for Length, predicted by NoEggs
```{r}
plot(NoEggs, Length, xlab = "NoEggs", ylab = "Length", col = ifelse(birdnest$Color == 1, "Red", "Black"), pch = 20)
# Add a title to the scatterplot
title(main = "Scatterplot of Length Predicted by NoEggs")
# Assign different colors in the scatterplot according to the condition of the variable “Color”
# Add a legend to distinguish the colors
legend("topright", legend = c("Colored", "Uncolored"), col = c("red", "black"), pch = 15)
```