---
title: "Four Step Process R Quiz (Amyloid Dataset)"
output:
  html_document:
    css: "../docs/style.css"
lang: en-US
---

### Load "Amyloid Altered" file
```{r}
amyloid <- read.csv("../data/amyloid.csv", header = TRUE)
attach(amyloid)
```

### Produce cooperative boxplots which show the Abeta values by Group
```{r}
boxplot(Abeta ~ Group, data = amyloid, main = "Abeta Values by Group", xlab = "Group", ylab = "Abeta Values", col = "lightblue") # nolint
```

### Produce cooperative dotplots which show the Abeta values by Group
```{r}
stripchart(Abeta ~ Group, data = amyloid, main = "Abeta Values by Group", xlab = "Group", ylab = "Abeta Values", col = "lightblue") # nolint
```

### Create a table for the n, mean, and sd for Abeta values by Group
```{r}
tapply(Abeta, Group, function(x) c(n = length(x), mean = mean(x), sd = sd(x)))
```

### Round table output to 2 decimal places
```{r}
tapply(Abeta, Group, function(x) c(n = length(x), mean = round(mean(x), 2), sd = round(sd(x), 2))) # nolint
```

### Run a t-test to see if there is a significant difference in Abeta values by Group # nolint
```{r}
t.test(Abeta ~ Group, data = amyloid)
```

### Give a one to two sentence interpretation of the t-test
```
Since the p value of 0.017 is less than 0.05, we reject Ho and we have convincing evidence that the true difference in means between group mAD and group NCI is not equal to 0
```

### Create a new column in the dataset for Residuals (Difference from observed values and Group mean) # nolint
```{r}
amyloid$residuals <- residuals(lm(Abeta ~ Group, data = amyloid))
```

### Check the Residuals with normal probability plots
```{r}
qqnorm(amyloid$residuals)
qqline(amyloid$residuals)
```

### Give a one to two sentence interpretation of the normal probability plots
```
The residuals are normally distributed because the points are close to the line. There may be possible outliers in the data past 2 theoretical quantiles.
```

### Find the Effect Size
```{r}
# Calculate the mean and standard deviation for each group
mean_nci <- mean(amyloid$Abeta[amyloid$Group == "NCI"])
mean_mad <- mean(amyloid$Abeta[amyloid$Group == "mAD"])
sd_nci <- sd(amyloid$Abeta[amyloid$Group == "NCI"])
sd_mad <- sd(amyloid$Abeta[amyloid$Group == "mAD"])

# Calculate the sample sizes for each group
n_nci <- length(amyloid$Abeta[amyloid$Group == "NCI"])
n_mad <- length(amyloid$Abeta[amyloid$Group == "mAD"])

# Compute the pooled standard deviation
pooled_sd <- sqrt(((n_nci - 1) * sd_nci^2 + (n_mad - 1) * sd_mad^2) / (n_nci + n_mad - 2)) # nolint

# Effect size
effect_size <- (mean_nci - mean_mad) / pooled_sd
effect_size
```