---
title: "Classwork Assignment 10/17/2024"
output:
  html_document:
    toc: true
    theme: darkly
    highlight: pygments
    code_download: true
lang: en-US
date: "Last Updated: `r format(Sys.Date(), '%B %d, %Y')` `r format(Sys.time(), '%H:%M:%S')`"
---

```{r, echo = FALSE}
brain <- read.csv("../data/BrainpH.csv", header = TRUE)
math <- read.csv("../data/MathEnrollment.csv", header = TRUE)
fluorescence <- read.csv("../data/Fluorescence.csv", header = TRUE)
```

## MathEnrollment Data Questions
### a. What percent of the variability in spring enrollment is explained by the multipe linear regression model based off fall enrollment and academic year?
```{r}
summary(lm(Spring ~ Fall + AYear, data = math))
```

`27.73%` of the variability in spring enrollment is explained by the multiple linear regression of fall enrollment and academic year.

### b. What is the size of the typical error for this multiple regression model?
```{r}
summary(lm(Spring ~ Fall + AYear, data = math))$sigma
```
The typical error for this multiple regression model is `31.08974`.

### c. Provide the ANOVA table for partitioning the total variability in spring enrollment based on this model and interpret the associated F-test.
```{r}
anova(lm(Spring ~ Fall + AYear, data = math))
```

The F-value for the "Fall" variable is `0.8763`, and the corresponding p-value is `0.3766`. Since the p-value is greater than the common significance level (e.g., 0.05), we fail to reject the null hypothesis. This indicates that the "Fall" variable does not have a statistically significant effect on spring enrollment.

 The F-value for the "AYear" variable is `2.1934`, with a p-value of `0.1769`. Again, since the p-value is greater than S.05, we fail to reject the null hypothesis, suggesting that "AYear" does not have a statistically significant effect on spring enrollment.

Based on the F-test results, neither "Fall" nor "AYear" significantly affects spring enrollment in this dataset, as indicated by the high p-values for both variables.

### d. Are the regression coefficients for both explanatory variables significantly different from zero? Provide appropriate hypotheses, test statistics, and P-values in order to make your conclusion.
```{r}
summary(lm(Spring ~ Fall + AYear, data = math))
```

Ho: The regression coefficient for "Fall" is zero.
Ha: The regression coefficient for "Fall" is not zero.

Fall: The estimated coefficient is -0.2021, with a standard error of 0.3589. The t-value is -0.563, and the p-value is 0.589.
AYear: The estimated coefficient is 4.5159, with a standard error of 3.0492. The t-value is 1.481, and the p-value is 0.177.

The decision rule is to reject the null hypothesis if the p-value is less than the significance level (commonly 0.05).

- For Fall, the p-value is 0.589, which is much greater than 0.05. Therefore, we fail to reject the null hypothesis. This suggests that the coefficient for "Fall" is not significantly different from zero.

- For AYear, the p-value is 0.177, which is also greater than 0.05. So, we also fail to reject the null hypothesis for "AYear," indicating that its coefficient is not significantly different from zero.

Based on the results, neither the "Fall" nor the "AYear" coefficients are significantly different from zero at the 0.05 significance level. This implies that neither variable has a statistically significant effect on the spring enrollment in this model.

## BrainpH Data Questions
### a. Make a scatterplot of `Y = pH` versus `X = Age and with Sex as a grouping variable (i.e., use different colors or different plotting symbols for the two levels of Sex). Comment on the plot.
```{r}
plot(brain$Age, brain$pH, col = ifelse(brain$Sex == "M", "blue", "red"), pch = 19, xlab = "Age", ylab = "pH")
legend("topright", legend = c("Male", "Female"), pch = 16, col = c("blue", "red"))
```

The scatterplot shows that the male pH levels are more spread out than the female pH levels.

### b. Fit the regression of pH on Age and test whether there is a linear association between the two variables.
```{r}
summary(lm(pH ~ Age, data = brain))
```

### c. Fit a model that produces parallel regression lines for the two levels of Sex. Write down the fitted prediction equation for each level of Sex.
```{r}
model_parallel <- lm(pH ~ Age + Sex, data = brain)
summary(model_parallel)
```

The prediction equation for Male is `pH = 6.9025758 - 0.0004535 * Age`

The prediction equation for Female is `pH = 6.9025758 - 0.0134258 * Age`

## Fluorescence Data Questions
### a. Fit a quadratic regression model for predicting ProteinProp from Calcium (if needed for the software you are using, create a new variable CalciumSq = Calcium x Calcium). Write down the fitted regression equation.
```{r}
fluorescence$CalciumSq <- fluorescence$Calcium^2
summary(lm(ProteinProp ~ Calcium + CalciumSq, data = fluorescence))
```

The fitted regression equation for Calcium is `ProteinProp = 0.479926 - 0.253189 * Calcium`

The fitted regression equation for CalciumSq is `ProteinProp = 0.479926 - 0.027788 * CalciumSq`

### b. Add the quadratic curve to a scatterplot of ProteinProp versus Calcium.
```{r}
model <- lm(ProteinProp ~ Calcium + CalciumSq, data = fluorescence)
plot(fluorescence$Calcium, fluorescence$ProteinProp, xlab = "Calcium", ylab = "ProteinProp", main = "Scatterplot with Quadratic Fit")
curve(coef(model)[1] + coef(model)[2] * x + coef(model)[3] * x^2, add = TRUE, col = "blue")
```


### c. Are the conditions for inference reasonably satisfied for this model?
```{r}
model <- lm(ProteinProp ~ Calcium + CalciumSq, data = fluorescence)
plot(fluorescence$Calcium, resid(model), xlab = "Calcium", ylab = "Residuals", main = "Residual Plot")
abline(h = 0, col = "red")
```

Not really, there's a clear pattern where the residuals start increasing and decreasing.

### d. Is the parameter for the quadratic term significantly different from zero? Justify your answer.
```{r}
summary(lm(ProteinProp ~ Calcium + CalciumSq, data = fluorescence))
```

### e. Identify the coefficient of multiple determination and interpret this value.
```{r}
summary(lm(ProteinProp ~ Calcium + CalciumSq, data = fluorescence))$r.squared
```