---
title: "Inference for Regression Slope Homework"
output:
  html_document:
    toc: true
    theme: darkly
    highlight: pygments
    code_download: true
lang: en-US
date: "Last Updated: `r format(Sys.Date(), '%B %d, %Y')` `r format(Sys.time(), '%H:%M:%S')`"
---

## Using Tests and Intervals

### Use AccordPrice file
```{r}
accord <- read.csv("../data/AccordPrice.csv", header = TRUE)
attach(accord)
```

### Run the linear model
```{r}
accord.lm <- lm(Price ~ Mileage, data = accord)
```

### Run a summary for the linear model
```{r}
summary(accord.lm)
```

### Compute 95% Confidence Intervals for coefficients
95% is the default. You can change the confidence level by changing the value of `conf.level`.
```{r}
confint(accord.lm, level = .95)
```

This agrees with our more manual computation on paper.

### To see the coefficients for either just the Intercepts, or just the slope
Remembering that .95 is the default level:
```{r}
confint(accord.lm, 1)
confint(accord.lm, 2)
```

### To find the Correlation Coefficient and Coefficient of Determination
```{r}
cor(Mileage, Price)
cor(Mileage, Price)^2
```

## Finding Prediction Intervals using R-Predict
Consider a accord with 50,000 miles. What is the 95% prediction interval for the price of this car?

> Note: Confidence Intervals speak to the placement of the mean, while Prediction Intervals speak to the placement of a specific value

> Given all other values being equal, Confidence Intervals will always be more narrow than Prediction Intervals

### First, find the 95% Confidence Interval
```{r}
newData <- data.frame(Mileage = 50)
predict(accord.lm, newData, interval = "confidence", level = .95)
```

### Next, find the 95% Prediction Interval
```{r}
predict(accord.lm, newData, interval = "prediction", level = .95)
```

#### Save this predict.plots function for use in later labs
```{r}
predict.plots <- function(x, y, conf.level = .95) {
  # x = explanatory variable
  # y = response variable
  model <- lm(y ~ x)
  new <- seq(min(x), max(x), length = 101)
  CI <- predict(model, list(x = new), int = "confidence", level = conf.level)
  PI <- predict(model, list(x = new), int = "prediction", level = conf.level)
  plot(x, y, ylim = range(y, PI[, 3]), main = "95% Confidence and Prediction Intervals", xlab = "Mileage", ylab = "Price")
  abline(model) # to obtain solid regression line
  points(new, CI[, 2], type = "l", col = "red", lty = 2)
  points(new, CI[, 3], type = "l", col = "red", lty = 2)
  points(new, PI[, 2], type = "l", col = "blue", lty = 3)
  points(new, PI[, 3], type = "l", col = "blue", lty = 3)
}
```

### Run predict.plots for our model
```{r}
predict.plots(Mileage, Price)
```