---
title: "R Markdown Quiz"
output:
  html_document:
    toc: true
    theme: darkly
    highlight: pygments
    code_download: true
lang: en-US
date: "Last Updated: `r format(Sys.Date(), '%B %d, %Y')` `r format(Sys.time(), '%H:%M:%S')`"
---

```{r, echo = FALSE}
baseball <- read.csv("../data/BaseballTimes.csv")
attach(baseball)
```

## Analyze the distribution of the response varuable (time in minutes) alone

###  Use and describe graphical displays (dotplot, histogram, boxplot)
```{r}
stripchart(Time, method = "stack", main = "Dotplot of Time", xlab = "Time (minutes)", ylab = "Frequency", pch = 19, col = "black")
hist(Time, main = "Histogram of Time", xlab = "Time (minutes)", col = "lightblue", breaks = 10)
boxplot(Time, main = "Boxplot of Time", xlab = "Time (minutes)", col = "lightblue", horizontal = TRUE)
```

###  Report descriptive statistics 
```{r}
summary(Time)
```

###  Identify the outliers using the 1.5 x IQR method 
```{r}
IQR <- IQR(Time)
outliers <- Time[Time < (quantile(Time)[2] - 1.5 * IQR) | Time > (quantile(Time)[4] + 1.5 * IQR)]
outliers
```

## Examine scatterplots to investigate which of the quantitative predictor variables appears to be the best single predictor of time.
```{r}
plot(Runs, Time, main = "Dotplot of Time vs Runs", xlab = "Runs", ylab = "Time (minutes)")
abline(lm(Time ~ Runs), col = "red")

plot(Margin, Time, main = "Dotplot of Time vs Margin", xlab = "Margin", ylab = "Time (minutes)")
abline(lm(Time ~ Margin), col = "red")

plot(Pitchers, Time, main = "Dotplot of Time vs Pitchers", xlab = "Pitchers", ylab = "Time (minutes)")
abline(lm(Time ~ Pitchers), col = "red")

plot(Attendance, Time, main = "Dotplot of Time vs Attendance", xlab = "Attendance", ylab = "Time (minutes)")
abline(lm(Time ~ Attendance), col = "red")
```

###  Comment on what the scatterplots reveal
- The scatterplot of `Time vs Runs` shows a positive linear relationship, suggesting that Runs may be a good predictor of Time, however, there appears to be a possible outlier at 15 Runs. 

- The scatterplot of `Time vs Margin` looks almost linear but with lots of variability, which doesn't look like a good predictor of Time.

- The scatterplot of `Time vs Pitchers` shows an almost perfect linear relationship with not much variability, making it the best predictor of Time among the variables examined.

- The scatterplot of `Time vs Attendance` shows no clear relationship between the two variables, suggesting that Attendance may not be a good predictor of Time.

###  Determine the regression equation for predicting time based on your chosen predictor
```{r}
time.lm <- lm(Time ~ Pitchers, data = baseball)
coef(time.lm)
```

###  Analyze conditions for inference including appropriate residual plots and comments on what they reveal
```{r}
par(mfrow = c(2, 2))
plot(time.lm, which = 1:4, main = "Residual Plots")
```

## Which game has the largest residual (in absolute value) for the model that you selected? 
```{r}
residuals <- residuals(time.lm)
max_residual <- which.max(abs(residuals))
baseball[max_residual, ]
```

###  Is this the same game that you identified as an outlier based on your analysis of the time variable alone?
Yes it is, the outlier we found from the time variable analysis was the same game with an outlier of 317.

## Repeat the entire analysis from the previous exercise, with the outlier omitted
```{r}
baseball_no_outlier <- baseball[-max_residual, ]
time.lm_no_outlier <- lm(Time ~ Pitchers, data = baseball_no_outlier)
coef(time.lm_no_outlier)
par(mfrow = c(2, 2))
plot(time.lm_no_outlier, which = 1:4, main = "Residual Plots (No Outlier)")
```

## Comment on the extent to which omitting the outlier changed the analysis and your conclusions.

Ommitting the outlier made the slope of the regression line steeper and the y-interval higher.