---
title: "Test 0916 (Gunnel and Placement Datasets)"
output:
  html_document:
    toc: true
    theme: darkly
    highlight: pygments
lang: en-US
date: "Last Updated: `r format(Sys.Date(), '%B %d, %Y')` `r format(Sys.time(), '%H:%M:%S')`"
---

## Part 1 

### Load and attach the Gunnels dataset
```{r}
gunnels <- read.csv("../data/gunnels.csv")
attach(gunnels)
```

### Find the 5-number summary for the variable Fromlow
```{r}
summary(Fromlow)
```

### Create a histogram for the variable Fromlow with breaks in multiples of 50
```{r}
hist(Fromlow, breaks = seq(0, 1000, by = 50), main = "Histogram of Fromlow", xlab = "Fromlow", ylab = "Frequency", col = "lightblue")
```

### Produce cooperative boxplots to compare the Fromlow variable across levels of the Gunnel variable
```{r}
boxplot(Fromlow ~ Gunnel, data = gunnels, main = "Boxplot of Fromlow by Gunnel", xlab = "Gunnel", ylab = "Fromlow", col = "lightblue")
```

### Create a table for the n, mean, and sd according to condition and round to two decimal places
```{r}
tapply(Fromlow, Gunnel, function(x) c(n = length(x), mean = round(mean(x), 2), sd = round(sd(x), 2)))
```

### Perform a t-test to evaluate the difference of Fromlow between the Gunnel groups
```{r}
t.test(Fromlow ~ Gunnel, data = gunnels)
```

### Add a column to your dataset for fitted values of Fromlow according to the Gunnel condition
```{r}
gunnels$fitted <- fitted(lm(Fromlow ~ Gunnel, data = gunnels))
```

### Create a qq plot and line for the Residuals for Gunnel Condition 0
```{r}
qqnorm(residuals(lm(Fromlow ~ Gunnel, data = gunnels)[gunnels$Gunnel == 0]), main = "QQ Plot of Residuals for Gunnel Condition 0", xlab = "Theoretical Quantiles", ylab = "Sample Quantiles")
qqline(residuals(lm(Fromlow ~ Gunnel, data = gunnels)[gunnels$Gunnel == 0]), col = "red", lwd = 2)
```

### Create a histogram for the Residuals for Gunnel Condition 0
```{r}
gunnels$residuals <- residuals(lm(Fromlow ~ Gunnel, data = gunnels))
hist(gunnels$residuals[gunnels$Gunnel == 0], main = "Histogram of Residuals for Gunnel Condition 0", xlab = "Residuals", ylab = "Frequency", col = "lightblue")
```

### Create a qq plot and line for the Residuals for Gunnel Condition 1
```{r}
qqnorm(gunnels$residuals[gunnels$Gunnel == 1], main = "QQ Plot of Residuals for Gunnel Condition 1", xlab = "Theoretical Quantiles", ylab = "Sample Quantiles")
qqline(gunnels$residuals[gunnels$Gunnel == 1], col = "red", lwd = 2)
```

### Create a histogram for the Residuals for Gunnel Condition 1
```{r}
hist(gunnels$residuals[gunnels$Gunnel == 1], main = "Histogram of Residuals for Gunnel Condition 1", xlab = "Residuals", ylab = "Frequency", col = "lightblue")
```

## Part 2

### Load and attach the Placement dataset
```{r}
placement <- read.csv("../data/placement.csv")
attach(placement)
```

### View the Scatterplot for cgpa according to placement_exam_marks
```{r}
plot(placement_exam_marks, cgpa, main = "Scatterplot of CGPA by Placement Exam Marks", xlab = "Placement Exam Marks", ylab = "CGPA")
```

### View the linear model
```{r}
lm(cgpa ~ placement_exam_marks, data = placement)
```

### View the summary for the linear model
```{r}
summary(lm(cgpa ~ placement_exam_marks, data = placement))
```

### Add the regression line to your scatterplot
```{r}
plot(placement_exam_marks, cgpa, main = "Scatterplot of CGPA by Placement Exam Marks", xlab = "Placement Exam Marks", ylab = "CGPA")
abline(lm(cgpa ~ placement_exam_marks, data = placement), col = "red", lwd = 2)
```

### Add a horizontal, dotted line at the mean cgpa
```{r}
plot(placement_exam_marks, cgpa, main = "Scatterplot of CGPA by Placement Exam Marks", xlab = "Placement Exam Marks", ylab = "CGPA")
abline(h = mean(cgpa), lty = 2, col = "red")
```

### Look at a plot of the residuals over fitted values
```{r}
plot(fitted(lm(cgpa ~ placement_exam_marks, data = placement)), residuals(lm(cgpa ~ placement_exam_marks, data = placement)), main = "Residuals vs Fitted Values", xlab = "Fitted Values", ylab = "Residuals")
```

### Look at the histogram of residuals
```{r}
hist(residuals(lm(cgpa ~ placement_exam_marks, data = placement)), main = "Histogram of Residuals", xlab = "Residuals", ylab = "Frequency", col = "lightblue")
```