---
title: "Outliers and Influential Points Notes"
output:
  html_document:
    toc: true
    theme: darkly
    highlight: pygments
lang: en-US
date: "Last Updated: `r format(Sys.Date(), '%B %d, %Y')` `r format(Sys.time(), '%H:%M:%S')`"
---

Analyze potential outliers in a dataset using residuals

## Introduction to standardized and studentized residuals
#### Load LongJumpOlympics dataset
```{r}
jump <- read.csv("../data/LongJumpOlympics.csv", header = TRUE)
attach(jump)
```

### Look at the plot of gold medal winning jumps, by year
```{r}
plot(Gold ~ Year, xlab = "Year", ylab = "Gold Medal Jump (meters)")
jump.lm <- lm(Gold ~ Year)
abline(jump.lm, col = "red")
```

Notice the 1968 performance of 8.9 meters. Is it an outlier?

### Look at the residuals
```{r}
jump.lm$residuals
plot(jump.lm$residuals ~ jump.lm$fitted.values, xlab = "Fitted Values", ylab = "Residuals", main = "Residuals vs. Fitted Values")
abline(h = 0, lty = 2, col = "red")
```

### What is the residual of the 1968 jump?
```{r}
jump.lm$residuals[16]
```

### Load the MASS library (necessary for standardized and studentized residuals)
```{r}
library(MASS)
```

### Look at the standardized residuals
```{r}
stdres(jump.lm)
plot(stdres(jump.lm) ~ jump.lm$fit, ylim = c(-4, 4), xlab = "Fitted Values", ylab = "Standardized Residuals", main = "Standardized Residuals vs. Fitted Values")
abline(h = -2, lty = 3, col = "red")
abline(h = 2, lty = 3, col = "red")
```

### The standardized residual of the 1968 jump was:
```{r}
stdres(jump.lm)[16]
```

...But the existence of the potential outlier affects the regression line...

Go back to the original plot of (Gold ~ Year)

### Add a red line to show what the regression line would look like without the 1968 data
```{r}
no68.df <- jump[!Year == "1968", ]
no68.lm <- lm(no68.df$Gold ~ no68.df$Year)
plot(Gold ~ Year, xlab = "Year", ylab = "Gold Medal Jump (meters)", main = "Gold Medal Winning Long Jumps")
abline(jump.lm)
abline(no68.lm, col = "red", lty = 3)
```

### What would be the standardized residual of the point if the regression line were computed without it (studentized residual)?
```{r}
stud_resids <- studres(jump.lm)
stud_resids
```

### Studentized Residual Plot
```{r}
plot(stud_resids ~ Year, ylim = c(-4, 4), xlab = "Year", ylab = "Studentized Residuals", main = "Studentized Residuals vs. Year")
abline(h = -2, lty = 3, col = "red")
abline(h = 2, lty = 3, col = "red")
```

### The studentized residual of the 1968 jump was
```{r}
stud_resids[16]
```

### New problem
```{r}
detach(jump)
```
## Analyze the PalmBeach dataset for outliers

### Load the PalmBeach dataset
```{r}
palm <- read.csv("../data/PalmBeach.csv", header = TRUE)
attach(palm)
```

### Scatter plot of the data
```{r}
plot(Buchanan, Bush, main = "Scatter Plot of Votes for Buchanan and Bush", xlab = "Votes for Buchanan", ylab = "Votes for Bush", pch = 19)
```

### Summary statistics for Buchanan
```{r}
summary(Buchanan)
```

### Summary statistics for Bush
```{r}
summary(Bush)
```

### Function to identify outliers using the IQR method
```{r}
identify_outliers <- function(x) {
  Q1 <- quantile(x, 0.25)
  Q3 <- quantile(x, 0.75)
  IQR <- Q3 - Q1
  lower_bound <- Q1 - 1.5 * IQR
  upper_bound <- Q3 + 1.5 * IQR
  outliers <- x[x < lower_bound | x > upper_bound]
  return(outliers)
}
```

### Outliers for Buchanan
```{r}
buchanan_outliers <- identify_outliers(Buchanan)
buchanan_outliers
```

### Outliers for Bush
```{r}
bush_outliers <- identify_outliers(Bush)
bush_outliers
```

### Interpretation
In the Palm Beach dataset, we found possible outliers by using the 1.5 IQR rule. There are possible outliers for Buchanan at `789, 650, 836, 3407, 1010`. For Bush, the possible outliers are at `177279, 152082, 176967, 152846, 184312`.