---
title: "Regression and Correlation Notes"
output:
  html_document:
    toc: true
    theme: darkly
    highlight: pygments
    code_download: true
lang: en-US
date: "Last Updated: `r format(Sys.Date(), '%B %d, %Y')` `r format(Sys.time(), '%H:%M:%S')`"
---

### Load the AccordPrice dataset and attach it
```{r}
accord <- read.csv("../data/AccordPrice.csv", header = TRUE)
accord
attach(accord)
```

### Find the mean of the Price column
```{r}
mean(Price)
```

### Find the linear model of price over milaege
```{r}
lm(Price ~ Mileage)
```

### Make a new column of expected values according to the linear model
```{r}
accord$Expected <- predict(lm(Price ~ Mileage))
accord$Mean <- mean(Price)
accord[, c("Expected", "Mean")]
attach(accord)
```

### Create ANOVA components
```{r}
accord$ModelSquares <- (Expected - Mean)^2
accord$ResidualSquares <- (Price - Expected)^2
accord$TotalSquares <- (Price - Mean)^2
accord[, c("ModelSquares", "ResidualSquares", "TotalSquares")]
```

### Sum the ANOVA components
```{r}
SSM <- sum(accord$ModelSquares)
SSE <- sum(accord$ResidualSquares)
SST <- sum(accord$TotalSquares)
c(SSM, SSE, SST)
```

### ANOVA Table
```{r}
anova(lm(Price ~ Mileage))
```

## Multiple methods to check linear relationship

### Confidence Interval
```{r}
confint(lm(Price ~ Mileage))
```

### T-test for slope
A t-test is the regression slope divided by the standard error of the regression
```{r}
summary(lm(Price ~ Mileage))
```
T-slope = `.1198 / .0141`

### T-test for correlation
```{r}
cor.test(Mileage, Price)
```

With simple linear regression, tstat for r and tstat slope are identical.

Also, F = t^2

Not the case with multiple predictors.