---
title: "Simple Linear Regression Notes"
output:
  html_document:
    toc: true
    theme: darkly
lang: en-US
date: "Last Updated: `r format(Sys.Date(), '%B %d, %Y')` `r format(Sys.time(), '%H:%M:%S')`"
---

```{r setup, echo = FALSE}
porsche <- read.csv("../data/PorschePrice.csv", header = TRUE)
```

### View the dataset
```{r}
porsche
```

### View the dimensions (rows x columns)
```{r}
dim(porsche)
```

### Check the column names
```{r}
colnames(porsche)
```

### Look at the first few records
```{r}
head(porsche, n = 3)
```
### Attach the file so that we can more easily refer to variable names
```{r}
attach(porsche)
```

### Look at a plot of mileage and price
```{r}
plot(Mileage, Price, main = "Price vs Mileage", xlab = "Mileage", ylab = "Price")
```

The data appear to be distributed in a roughly linear configuration with a negative trend

Use R's lm function to see the linear model (the defined equation for y-hat)

Store the linear model as porsche_lm
```{r}
porsche_lm <- lm(Price ~ Mileage)
```

The lm function returns a list with a lot of useful information
```{r}
length(porsche_lm)
names(porsche_lm)
```

### View the linear model
```{r}
porsche_lm
```

### Other ways to see information
```{r}
porsche_lm$coefficients
porsche_lm$coef
```

This tells us that the line of best fit for our sample data is pricehat=71.0905 - .5894(Mileage)

### View the regression line
```{r}
plot(Mileage, Price, main = "Price vs Mileage", xlab = "Mileage", ylab = "Price")
abline(porsche_lm, )
```

#### You can also write abline by giving any intercept and slope such as
```{r}
plot(Mileage, Price, main = "Price vs Mileage", xlab = "Mileage", ylab = "Price")
abline(71.0905, -.5894)
```

### Add a horizontal, dotted line at the mean price level
```{r}
plot(Mileage, Price, main = "Price vs. Mileage", xlab = "Mileage", ylab = "Price")
abline(h = mean(Price), lty = 3, col = "red")
```

#### lty gives characteristics of the line (line type)
- lty1 = solid
- lty2 = dashed
- lty3 = dotted
- lty4 = dot-dash
- lty5 = long dash

### Look at the fitted values and residuals
```{r}
porsche_lm$residuals
cbind(Mileage, Price, porsche_lm$fitted.values, porsche_lm$residuals)
```

### Do the same thing, but only look at the first 5 rows
```{r}
cbind(Mileage, Price, porsche_lm$fitted.values, porsche_lm$residuals)[1:5, ]
```

### Or use the head function to look at the top of the dataset
```{r}
head(cbind(Mileage, Price, porsche_lm$fitted.values, porsche_lm$residuals))
```

### Save the table
```{r}
porscheTable <- data.frame(Mileage, Price, fits = porsche_lm$fitted.values, resids = porsche_lm$residuals)
```

### View column names
```{r}
colnames(porscheTable)
```

### Change the names of columns 3 and 4
```{r}
names(porscheTable)[c(3, 4)] = c("fits", "resids")
```

### Check your work
```{r}
str(porscheTable)
porscheTable
```

### Summary
```{r}
summary(porsche_lm)
```